Characterizing Stellar Streams by ML

### 2023-08-31 src/tex/ms.tex
[showyourwork] overleaf sync

        The custom model framework combines for each model a neural network with
        Bayesian probability functions. The latter are set by the model type,
        e.g. the Gaussian model has a Gaussian likelihood function, but the
        former, the neural network, is set and determined by the user. So long
        as the network has the correct number of inputs (e.g. one for $\phi_1$)
        and outputs (one per parameter), any feed-forward network may be used.
        Our default choice, and for which we provide a helper function, is an
        MLP with a sequence of \texttt{Linear} \!\!\textrightarrow \texttt{Tanh}
        \!\!\textrightarrow \texttt{Dropout} (hereafter \texttt{LinTanh})
        blocks. This is illustrated in \autoref{table:mlp_structure}.
        % FIXME! reference not being picked up correctly.
        Dropout layers are used to prevent network co-adaptation and
        over-fitting. Dropout works by randomly zeroing out nodes in the
        network, decreasing the importance of any single node, encouraging
        instead emergent patterns across the network.  One model for which the
        lin-tanh-dropout network cannot be used is the normalizing flow used in
        the photometric background model
        (\autoref{sub:method:photometric_model:off_stream}). Here we use the
        \package{nflows} package, with a conditional diagonal normal base
        distribution and composite transform network of a reverse permutation
        and masked affine auto-regressive transform layers. As noted in our data
        availability statement, all the codes used in this paper are bundled by
        \href{https://github.com/showyourwork/showyourwork}{\package{showyourwork}}
        in a public repository at $\GitHubURL{}$. We will describe for each
        stream model specifics relevant to that stream but refer readers to the
        publicly-available code itself for code implementation details.
        % TODO: Maybe more details in the Appendix?

@@

        We utilize data from \Gaia{} DR3 for the astrometric coordinates of each
        star.  To obtain more accurate color information, we cross-match the
        \Gaia{} field on \PanStarrs{} \citep{PanSTARRS1} using \Gaia's provided
        \texttt{Best-neighbors} catalog.  To limit the field to a region known
        to contain \stream{GD-1} we adopt broad cuts on the cross-matched data.
        We perform these rough cuts by requiring the data lie within the
        phase-space cube:
        \begin{itemize}
            \item $\phi_1 \in [-100, 40] \ \rm{\degree}$,
            \item $\phi_2 \in [-9, 5] \rm{\degree}$,
            \item $\varpi \in [-10, 1] \ \rm{mas}$,
            \item $\rm{bp-rp} \in [-1, 3] \ \rm{mag}$
            \item $g, r \in [0, 50] \ \rm{mag}$, and not \texttt{null}.
        \end{itemize}
        where, we use \package{gala}'s \citep{gala, galav1.3}
        \code{GD1Koposov10} frame \citep{Koposov+2010} to define $\phi_1$,
        $\phi_2$, and associated proper motions.  For downloading purposes, we
        split the query into 10 sub-regions, in even $\phi_1$ increments. A plot
        of the $\phi_1, \phi_2$ regions is included in the
        \href{\GitHubURL{}}{\package{ShowYourWork}} repository.  Stream-specific
        coordinates ($\phi_1, \phi_2$) are not included in \Gaia.  One could
        embed the coordinate transform in the ADQL \citep{ADQL2.0}, or define
        great-circle arcs in ICRS \citep{ICRS1997}. Since only rough cuts are
        required and since the sub-region boxes are sufficiently small for a
        reasonable flat-sky approximation  we simply translate the ($\phi_1,
        \phi_2$) boxes to ICRS.  We correct the \PanStarrs{} photometry for dust
        extinction using the \textit{Bayestar19} \citep{Green+2019} models in
        the \package{dustmaps} package \citep{Green2018}.  \textit{Bayestar19}
        is a three-dimensional dust map of the Milky Way for which we use
        \Gaia's parallax distance information.  We include the $1-\sigma$ errors
        as a source of uncertainty in $\mbs{\Sigma}_n$, the per-star photometric
        uncertainty.
        % TODO: actually implement that in the model
        The result is a field populated by
        \variable{output/gd1/ndata_variable.txt} stars, from which we would like
        to identify stream members.

>>>>

        The custom model framework combines for each model a neural network with
        Bayesian probability functions. The latter are set by the model type,
        e.g. the Gaussian model has a Gaussian likelihood function, but the
        former, the neural network, is set and determined by the user. So long
        as the network has the correct number of inputs (e.g. one for $\phi_1$)
        and outputs (one per parameter), any feed-forward network may be used.
        Our default choice, and for which we provide a helper function, is an
        MLP with a sequence of \texttt{Linear} \!\!\textrightarrow \texttt{Tanh}
        \!\!\textrightarrow \texttt{Dropout} (hereafter \texttt{LinTanh})
        blocks. This is illustrated in \autoref{table:mlp_structure}.
        % FIXME! reference not being picked up correctly.
        Dropout layers are used to prevent network co-adaptation and
        over-fitting. Dropout works by randomly zeroing out nodes in the
        network, decreasing the importance of any single node, encouraging
        instead emergent patterns across the network.  One model for which the
        lin-tanh-dropout network cannot be used is the normalizing flow used in
        the photometric background model
        (\autoref{sub:method:photometric_model:off_stream}). Here we use the
        \package{nflows} package, with a conditional diagonal normal base
        distribution and composite transform network of a reverse permutation
        and masked affine auto-regressive transform layers. As noted in our data
        availability statement, all the codes used in this paper are bundled by
        \href{https://github.com/showyourwork/showyourwork}{\package{showyourwork}}
        in a public repository at \GitHubURL{}. We will describe for each
        stream model specifics relevant to that stream but refer readers to the
        publicly-available code itself for code implementation details.
        % TODO: Maybe more details in the Appendix?

@@

        We utilize data from \Gaia{} DR3 for the astrometric coordinates of each
        star.  To obtain more accurate color information, we cross-match the
        \Gaia{} field on \PanStarrs{} \citep{PanSTARRS1} using \Gaia's provided
        \texttt{Best-neighbors} catalog.  To limit the field to a region known
        to contain \stream{GD-1} we adopt broad cuts on the cross-matched data.
        We perform these rough cuts by requiring the data lie within the
        phase-space cube:
        \begin{itemize}
            \item $\phi_1 \in [-100, 40] \ \rm{\degree}$,
            \item $\phi_2 \in [-9, 5] \rm{\degree}$,
            \item $\varpi \in [-10, 1] \ \rm{mas}$,
            \item $\rm{bp-rp} \in [-1, 3] \ \rm{mag}$
            \item $g, r \in [0, 50] \ \rm{mag}$, and not \texttt{null}.
        \end{itemize}
        where we use \package{gala}'s \citep{gala, galav1.3}
        \code{GD1Koposov10} frame \citep{Koposov+2010} to define $\phi_1$,
        $\phi_2$, and associated proper motions.  For downloading purposes we
        split the query into 10 sub-regions in even $\phi_1$ increments. A plot
        of the $\phi_1, \phi_2$ regions is included in the
        \href{\GitHubURL{}}{\package{ShowYourWork}} repository.  Stream-specific
        coordinates ($\phi_1, \phi_2$) are not included in \Gaia.  One could
        embed the coordinate transform in the ADQL \citep{ADQL2.0}, or define
        great-circle arcs in ICRS \citep{ICRS1997}. Since only rough cuts are
        required and since the sub-region boxes are sufficiently small for a
        reasonable flat-sky approximation  we simply translate the ($\phi_1,
        \phi_2$) boxes to ICRS.  We correct the \PanStarrs{} photometry for dust
        extinction using the \textit{Bayestar19} \citep{Green+2019} models in
        the \package{dustmaps} package \citep{Green2018}.  \textit{Bayestar19}
        is a three-dimensional dust map of the Milky Way for which we use
        \Gaia's parallax distance information.  We include the $1-\sigma$ errors
        as a source of uncertainty in $\mbs{\Sigma}_n$, the per-star photometric
        uncertainty.
        % TODO: actually implement that in the model
        The result is a field populated by
        \variable{output/gd1/ndata_variable.txt} stars, from which we would like
        to identify stream members.
